### 2019-11-08 fertile.tex
abstract - draft #1

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Title here}

  \author{
        Author 1 \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of YYY, University of XXX\\
     and \\     Author 2 \\
    Department of ZZZ, University of WWW\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title here}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The text of your abstract. 200 or fewer words.
\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\section{Introduction}\label{introduction}

\section{The Importance of Ensuring
Reproducibility}\label{the-importance-of-ensuring-reproducibility}

Practicing reproducible research is key for advancing our scientific
knowledge. One reason for this is that reproducible research encourages
transparency. Providing one's audience with well organized, reproducible
code and results allows readers to understand the steps taken to
generate findings from raw data and determine for themselves whether or
not they believe the results to be reliable or trustworthy.

Additionally, reproducibility encourages collaboration and extended
research, allows others to easily apply the methods used in one project
to their own work with minimal effort. \citet{bray2014five}

Fertile is designed to make reproducibility simple, providing fast and
easy methods to test an R project for reproducibility. The package is
intended to be usable by introductory data science students in their
first semester of R.

--\textgreater{} make sure this info is included somewhere:

From ``Five Concrete Reasons Your Students Should Be Learning to Analyze
Data in the Reproducible Paradigm'' (\citet{bray2014five})

\begin{itemize}
\tightlist
\item
  Helps with transparency: make clear data cleaning steps between raw
  data and final data
\item
  Helps with collaboration: easier to share code when it takes very few
  steps to run on a different computer
\end{itemize}

The other big ideas on the importance/use of this package (not from
sources):

\begin{itemize}
\tightlist
\item
  Make it easier for professors to grade students' code
\item
  Should be usable by intro level data science students
\item
  Possible use by reviewers of journal articles (and by those writing
  the articles)
\end{itemize}

\section{What Defines Reproducibility in Data
Science?}\label{what-defines-reproducibility-in-data-science}

\section{\texorpdfstring{Creating Comprehensive Reproducibility Reports
With
``fertile''}{Creating Comprehensive Reproducibility Reports With fertile}}\label{creating-comprehensive-reproducibility-reports-with-fertile}

Main points of different sources, as well as info about how they might
be used for the paper.

Here are some sources we might use for motivation behind the project:

\subsubsection{The Reproducibility
Crisis}\label{the-reproducibility-crisis}

Big idea: most scientific fields are facing a reproducibility crisis and
poor statistical use is considered one of the important reasons behind
this.

Not sure how useful this is due to the fact that it does not necessarily
focus on the same kind of reproducibility we are looking at, which is
code reproducibility rather than experimental reproducibility.

\citet{baker20161}

\subsubsection{Why is reproducibility
important?}\label{why-is-reproducibility-important}

From Popper in \texttt{The\ Logic\ of\ Scientific\ Discovery}:
``non-reproducible single occurrences are of no significance to
science.''

\citet{popper2005logic}

\subsubsection{What does reproducibility mean in data
science?}\label{what-does-reproducibility-mean-in-data-science}

``The ability of a researcher to duplicate the results of a prior study
using the same materials as were used by the original investigator. That
is, a second researcher might use the same raw data to build the same
analysis files and implement the same statistical analysis in an attempt
to yield the same results.''

\citet{goodman2016does}

From ``The Practice of Reproducibile Research'': Can all of the
figures/calculations related to the result be reproduced in a single
button press or at least with a reasonably small effort?

\citet{kitzes2017practice}

\subsubsection{What makes a data science project
reproducibile?}\label{what-makes-a-data-science-project-reproducibile}

Another way to think about this is: what features does fertile have that
match what different sources think needs to be checked to ensure
reproducibility?

Some big ideas from the OpenSci discussion page:

\begin{itemize}
\tightlist
\item
  code should use tidy style
\item
  project should have short vignette files -- most written code should
  be in an R/ directory
\item
  code and data should be stored in separate folders
\item
  a run environment and dependencies should be specified
\item
  there should be a README file
\item
  there should be a data dictionary
\item
  one master script
\item
  code should only use relative paths
\item
  code should be well commented and all variables should be defined
\end{itemize}

Big ideas from ``Packaging data analytical work reproducibly using R''

\begin{itemize}
\tightlist
\item
  research should be organized like an R package!
\item
  clear separation of data, method, and output
\item
  specify the computational environment that was used for the original
  analysis (typically in a plain text file)
\item
  there should be a README describing the project and where to get
  started
\item
  script files with reusable functions should go in an R/ directory
\item
  raw data files should be kept in a data/ directory
\item
  analysis scripts and report files should go in an analysis/ directory
\item
  for simpe projects, scripts should be given ascending names. For more
  complicated ones, the use of some sort of makefile is recommended
\item
  there should also be a DESCRIPTION file with information about the
  authors, project license, and software dependencies
\end{itemize}

\citet{marwick2018packaging}

From ``The Practice of Reproducible Research''

\begin{itemize}
\item
  Are the data openly accessible? If hosted online, is the web address
  reliable long-term?
\item
  Are they in a commonly used and well-documented file format? Avoid
  spreadsheets and instead use plain text data if possible!
\item
  Is the raw data available? Is sufficient metadata provided?
\item
  Are dependencies described properly?
\item
  Is full history of source code available through a public version
  history
\item
  Is there a README?
\item
  Are functions documented?
\item
  Is there narrative documentation explaining how the different pieces
  work together?
\item
  Are there usage examples?
\end{itemize}

Folder setup might look like:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Raw Data

  \begin{itemize}
  \tightlist
  \item
    Data
  \item
    README
  \end{itemize}
\item
  Clean Data

  \begin{itemize}
  \tightlist
  \item
    Data
  \end{itemize}
\item
  Results

  \begin{itemize}
  \tightlist
  \item
    Results file
  \end{itemize}
\item
  Src

  \begin{itemize}
  \tightlist
  \item
    Analysis script
  \item
    Script to clean data
  \end{itemize}
\end{enumerate}

\citet{kitzes2017practice}

From R OpenSci's Reproducibility Guide:

\url{http://ropensci.github.io/reproducibility-guide/}

\begin{itemize}
\tightlist
\item
  Is it clear where to begin?
\item
  Can you determinine which files were used as input to create output
  files?
\item
  Is there documentation about every result?
\item
  Are exact versions of external applications noted?
\item
  If using randomness, are seeds noted?
\item
  Have you specified a license or noted licenses if you used other
  people's content?
\item
  Are files easy to find?
\item
  Is it clear what the most recent file is?
\item
  Are there any folders that could be deleted?
\item
  Is analysis output done hierarchically?
\item
  Are there lots of manual data manipulation steps?
\end{itemize}

From ``A Guide to Reproducible Code in Ecology and Evolution'' (these
ideas are pretty universal, though):

A basic project structure:

\begin{itemize}
\item
  The data folder contains all input data (and metadata) used in the
  analysis.
\item
  The doc folder contains the manuscript
\item
  The figs directory contains figures generated by the analysis
\item
  The output folder contains any type of intermediate or output files
  (e.g.~simulation outputs, models, processed datasets, etc.). You might
  separate this and also have a cleaned-data folder.
\item
  The R directory contains R scripts with function definitions.
\item
  The reports folder contains RMarkdown files that document the analysis
  or report on results
\item
  Consistent, ordered naming of scripts
\item
  Use portable paths
\item
  Write unit tests (only for advanced coding)
\item
  Show the packages you used
\item
  Record dependencies and versions of outside things you use
\end{itemize}

\citet{cooper2017guide}

\section{Karl Broman's Suggestions}\label{karl-bromans-suggestions}

\begin{itemize}
\tightlist
\item
  Encapsulate everything within one directory
\item
  Separate raw data from derived data
\item
  Separate data from code
\item
  Use relative paths
\item
  Choose filenames correctly
\item
  Write README files
\end{itemize}

\url{http://kbroman.org/steps2rr/pages/organize.html}

\section{Why focus on R?}\label{why-focus-on-r}

\begin{itemize}
\tightlist
\item
  R is the most popular language for statistical programming and is
  specifically designed for statistics
\item
  R is great for reprodubility because the code is readable by users,
  and RMarkdown is a great way to show/explain processes
\item
  R is easy to install and begin using
\end{itemize}

--\textgreater{} from
\url{https://openresearchsoftware.metajnl.com/articles/10.5334/jors.bu/print/}

\section{Similar R Packages}\label{similar-r-packages}

\subsubsection{rrtools}\label{rrtools}

\url{https://github.com/benmarwick/rrtools}

\begin{itemize}
\tightlist
\item
  Creates a basic R package named after your research topic
\item
  Generates a license file
\item
  Connects to GitHub and creates a repository
\item
  Generates a README
\item
  Generates a reproducible directory structure
\item
  Creates a dockerfile
\item
  Creates a minimal travis file
\item
  Sets up testthat
\end{itemize}

\section{Msc sources to look at:}\label{msc-sources-to-look-at}

\subsubsection{\texorpdfstring{Victoria Stodden's ``Implementing
Reproducible Research''
book}{Victoria Stodden's Implementing Reproducible Research book}}\label{victoria-stoddens-implementing-reproducible-research-book}

\url{https://books.google.com/books?hl=en\&lr}=\&id=JcmSAwAAQBAJ\&oi=fnd\&pg=PP1\&dq=Victoria+stodden+reproducibility\&ots=ym\_gvSxMNH\&sig=B5N6FlkoocVjIprIZevtvrL33qQ\#v=onepage\&q=coombes\&f=false

\bibliographystyle{agsm}
\bibliography{bibliography.bib}

\end{document}