### 2020-01-13 fertile_sdss_presentation.Rmd
presentation files

---
title: "Creating optimal conditions for reproducible data analysis in R with 'fertile'"
subtitle: 
author: |
  Audrey Bertin<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  `r icon::fa("envelope")` &nbsp;[abertin@smith.edu](mailto:abertin@smith.edu) | 
  `r icon::fa("twitter")` @[ambertin99](http://www.twitter.com/ambertin99) |
  `r icon::fa("github")` [ambertin](http://www.github.com/ambertin)
  </span>
date: |
  SDSS 2020 <br/>
  Date <br/>
  `r sds::img_logo(width = 64)`
output:
  sds::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
# sds::duo_smith()
sds::mono_light_smith()
```

# Reproducibility

.pull-left[

• When code + data files produce **identical** results when run by another analyst

### Why this is important:
  - Research is increasingly data-driven
  - Knowledge can be shared worldwide rapidly
  - Transparency
  - Specific feedback
  - Sharing of knowledge / future work

]

.pull-right[

![Diagram by Stephanie Hicks and Roger Peng](images/reproducibility.png)

]

---

# Reproducibility challenges

• No consensus on standards of reproducibility in research
  - **components** (e.g. need it be open source?)
  - **recommendations** (e.g. should raw / processed data be stored separately?)

• Discussion is very generalized
  - For any software
  - Difficult to implement (particularly for new analysts)

---

# fertile addresses this

• `fertile`: An R package that focuses on common mistakes data analysts make in R

### Two modes

• Proactive 
  - Stops mistakes as they are happening

• Retroactive
  - Analyzes already-written code for potential issues

---
# Proactive use

• Identifies potential mistakes as they are made

• Outputs an informative message and a recommended solution

.pull-left[

• Ex: detecting problematic file paths (absolute or pointing outside a directory)

![](images/absolute.png)

]

.pull-right[

• Ex: setting working directories

![](images/setwd.png)

]

---
# Retroactive use

• Analyze an RStudio project:
  - Directory structure
  - File path use
  - Randomness

• Provide recommendations for remedies

• Features easily accessed w/ two simple functions

---
# Retroactive use

### 1. `proj_check()`

• Runs 15 reproducibility checks
  - Clear build chain
  - Clutter-free project root level
  - No unused files
  - Randomness without `set.seed()`

• Notes which pass/fail

• Gives recommendations for addressing issues

• Gives a guide of where to look for help

---

![](images/proj_check.png)

---
# Retroactive use

### 2. `proj_analyze()`

• Creates a report focused primarily on the structure of a data analysis project

• Contains:
  - Packages referenced
  - Files / types
  - Suggestions for reorganization
  - Problematic file paths

---

![](images/proj_analyze.png)

---
# Why `fertile` matters

• `fertile` lowers barriers to reproducibile data analysis in R

• Provides non-generalized recommendations

• Near-effortless accessibility

• Good for analysts of all skill levels / backgrounds

--- 
---

# `fertile` is different

• There are already R packages (some still in development) trying to address this issue:
  -rrtools
  -represtools
  -orderly
  -workflowr
  -checkers

• However, many of these are:
  -Complicated and difficult for beginners
  -Focused on just one aspect of reproducibility

• `fertile` is incredibly simple to use and covers a wide variety of topics

---

# Testing `fertile`

• Is `fertile` actually effective for teaching reproducibility?
• Experimented in a Smith College introductory data science course

• Findings:

---

class: center, middle
# Thank you!

---
class: center, middle
## Acknowledgements

The authors gratefully acknowledge contributions from:
Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the `tidyverse` team.

---

`r icon::fa("envelope")` [abertin@smith.edu](mailto:abertin@smith.edu) |
`r icon::fa("twitter")` [ambertin99](http://www.twitter.com/ambertin99) |
`r icon::fa("github")` [ambertin](http://www.github.com/ambertin)